A user-extensible and adaptable parser architecture

### 2008-05-30 doc/logparser-acronyms.tex
Improved coverage section - it was better than I had thought. Rewrote parts of the dips in the graph explanation. Minor changes to shorten the paper.

\newacronym{DNSBL}{DNS Blacklist}{description={
    A DNS Blacklist~\cite{Wikipedia-DNSBL} is a simple collaborative
    anti-spam technique used to reject or penalise email sent from mail
    servers reported to have sent large volumes of spam.  To use a DNSBL
    the mail server makes a DNS request incorporating the IP address of the
    currently connected client; if the requested hostname exists the client
    is on the DNSBL, and the mail server can decide what course of action
    to take.
}}

@@

\newacronym{SLCT}{Simple Logfile Clustering Tool}{description={
    SLCT~\cite{slct-paper} is a tool implementing an algorithm designed by
    Risto Vaarandi for identifying and grouping similar log files, then
    producing a regex for each group of lines.
}}

>>>>

\newacronym{DNSBL}{DNS Blacklist}{description={
    A DNS Blacklist~\cite{Wikipedia-DNSBL} is a simple collaborative
    anti-spam technique used to reject or penalise email sent from mail
    servers reported to have sent large volumes of spam.  To use a DNSBL
    the mail server makes a DNS request incorporating the IP address of the
    client; if the requested hostname exists the client is on the DNSBL,
    and the mail server can decide what course of action to take.
}}

@@

\newacronym{SLCT}{Simple Logfile Clustering Tool}{description={
    SLCT~\cite{slct-paper} is a tool implementing an algorithm designed by
    Risto Vaarandi for identifying, grouping and producing a regex to match
    similar log lines.
}}

### 2008-05-30 papers/sgai-2008/paper.tex
Improved coverage section - it was better than I had thought. Rewrote parts of the dips in the graph explanation. Minor changes to shorten the paper.

XXX MINIMUM VERSUS MAXIMUM NUMBER OF RULES\@.

@@

The dips at log files 22 and 62--68 in fig.~\refwithpage{percentage
increase of shuffled and reversed over normal} correspond to peaks in log
file size in fig.~\refwithpage{execution time vs file size vs number of
lines graph}, and peaks in fig.~\refwithpage{execution time vs file size vs
number lines factor} (where a peak means that more lines are processed per
second, i.e.\ performance is better).  The explanation for this took some
time to arrive at, but it turns out to be reasonably simple.  The large log
files in question were caused by a mail forwarding loop, where the
distribution of log lines is quite different to normal, resulting in
different performance characteristics.

XXX REWRITE THIS NEXT PARAGRAPH SO IT DOESN'T MENTION postfix/smtpd.  The
vast majority of log lines when a mail loop occurs are from Postfix
components which have a small number of rules associated with them, whereas
in general \daemon{smtpd} adds the majority of log lines, and also has the
highest number of rules.  \daemon{smtpd} log lines are distributed across
rules much more evenly than the log lines which occur frequently during a
mail loop, so the average number of rules required to parse a
\daemon{smtpd} log line is much higher that the average number required to
parse other log lines.

These two characteristics combine to reduce the average number or rules
required to parse a log line when there is a mail loop, as shown by the
peaks in fig.~\refwithpage{execution time vs file size vs number lines
factor} and troughs in fig.~\refwithpage{percentage increase of shuffled
and reversed over normal}.  When the rule ordering is reversed the majority
of log lines generated by a mail loop will be parsed using very few rules,
because there are very few rules for the Postfix components producing the
log lines.  Without a mail loop the Postfix components producing the
majority of log lines have a large number of rules associated with them,
and reverse ordering deliberately starts with the least likely rules and
moves towards the most likely.  This leads to a noticeable drop in the
average time required to parse a log line during a mail loop, as shown in
fig.~\refwithpage{percentage increase of shuffled and reversed over
normal}.  The number of rules which need to be consulted when the ordering
is shuffled varies between the optimum and the nadir, and the performance
varies proportionally.

@@

XXX THIS NEEDS TO BE IMPROVED AND PROOFREAD\@.

The discussion of the parser's coverage of Postfix log files is separated
into two parts: log lines covered and mails covered.  The first is
important because the parser should handle all (relevant) log lines it's
given; the second is equally important because the parser must properly
deal with every mail if it is to be useful.  Improving the former is less
intrusive, as it just requires new rules to be written; improving the
latter is much more intrusive as it requires changes to the parser
algorithm, and it can also be much harder to notice a deficiency.

Warnings are issued for any log lines which are not parsed; no warnings are
issued for unparsed log lines while testing with the \numberOFlogFILES{}
test log files, so it can be safely concluded that there are zero false
negatives.  False positives are harder to quantify: short of examining each
of the 60,721,709 log lines and determining which \regex{} parsed it, there
is no way to be sure that every line was parsed by the correct \regex{},
making it impossible to quantify the false positive rate; however a random
sample of 6039 log lines was parsed and the results checked manually to
ensure that the correct \regex{} parsed each log line.

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails which were not handled properly.  The parser
issues warnings when it detects any errors, some of which may alert the
user to a problem.  There should be few or no warnings when parsing, and
when finished parsing the state table should only contain entries for mails
which had yet to be delivered when the log files ended, or were accepted
before the log files began.

Verifying by inspection that the parser correctly deals with all 60,721,709
lines in the test log files is infeasible, but verifying a subset of those
log files is a tractable, if extremely time consuming, task.  A sample of
log lines was parsed with all debugging options enabled, resulting in
167,448 lines of output.  All 167,448 lines were examined in conjunction
with the input log and a dump of the resulting database, verifying that for
each of the input lines the parser used the correct rule and executed the
correct action, which in turn produced the correct result and inserted the
correct data in the database.  The log segment produced 4 warnings, 10
mails remaining in the state tables, and 1625 connections correctly entered
in the database.

Given the evidence detailed above, the author is confident that the false
positive rate when reconstructing a mail is exceedingly low, if not
approaching zero.

Parser coverage is divided into two topics in this section: log lines
covered and mails covered.  The former is initially more important, as the
parser must successfully parse every line if it is to be complete, but
subsequently the latter takes precedence because reproducing the path a
mail takes through Postfix is the aim of the parser.  Increasing the
percentage of log lines parsed is relatively simple and non-intrusive:
adding new rules or modifying existing rules is simplified by the
separation of rules, actions and framework.  Improving the logical coverage
is harder, as the actions taken by Postfix must be reconstructed by the
author, and the new sequence of actions integrated into the existing model
without breaking the existing parsing.  Detecting a deficiency in the
parsing algorithm is also significantly harder than detecting unparsed log
lines, as the parser will warn about any unparsed line, whereas discovering
a flaw in the parser requires understanding of the warnings produced and
the mails remaining in the state table.  Rectifying a flaw in the parser
requires an understanding of both the parser and Postfix's log files, and
investigative work to determine the cause of the deficiency, followed by
further examination of the log files in developing a solution.

>>>>

@@

The dips at log files 22 and 62--68 in fig.~\refwithpage{percentage
increase of shuffled and reversed over normal} correspond to peaks in log
file size in fig.~\refwithpage{execution time vs file size vs number of
lines graph}, and peaks in fig.~\refwithpage{execution time vs file size vs
number lines factor} (where a peak means that performance is better).  The
explanation for this took some time to arrive at, but it turns out to be
reasonably simple.  The large log files in question were caused by a mail
forwarding loop, where the distribution of log lines is quite different to
normal, resulting in different performance characteristics.  The vast
majority of log lines when a mail loop occurs are from Postfix components
which have a small number of rules associated with them, whereas in general
the majority of log lines are from the Postfix component (\daemon{smtpd})
with the highest number of rules.  Furthermore in normal log files the
majority of log lines are distributed much more evenly across rules than
the majority of log lines during a mail loop.

These two characteristics combine to reduce the average number or rules
required to parse a log line when there is a mail loop, as shown by the
peaks in fig.~\refwithpage{execution time vs file size vs number lines
factor} and troughs in fig.~\refwithpage{percentage increase of shuffled
and reversed over normal}.  When the rule ordering is reversed the majority
of log lines generated by a mail loop will be parsed using very few rules,
because there are very few rules for the Postfix components producing the
log lines.  Without a mail loop the Postfix components producing the
majority of log lines have a large number of rules associated with them,
and reverse ordering deliberately starts with the least likely rules and
moves towards the most likely.  This leads to a noticeable drop in the
average time required to parse a log line during a mail loop, as shown in
fig.~\refwithpage{percentage increase of shuffled and reversed over
normal}.  The number of rules which need to be consulted when the ordering
is shuffled varies between the optimal and the pessimal, with the
performance varying proportionally.

@@

The parser's coverage of Postfix log files is separated into two parts: log
line coverage and mail coverage.  The first is important because the parser
should handle all input log lines; the second is equally important because
the parser must properly reconstruct every mail if it is to be useful.
Improving the former is less intrusive, as it just requires new rules to be
written; improving the latter is more intrusive as it requires changes to
actions, and it can also be much harder to notice a deficiency.

The framework issues warnings for any log lines which are not parsed; no
warnings are issued for unparsed log lines while parsing the
\numberOFlogFILES{} test log files, so it can be safely concluded that
there are zero false negatives.  False positives are harder to quantify:
manually verifying that the correct \regex{} parsed each of the 60,721,709
log lines is infeasible, making it impossible to quantify the false
positive rate; however a random sample of 6039 log lines (0.00994\% of the
total) was parsed and the results verified to ensure that the correct
\regex{} parsed each log line.

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails which were not handled properly.  The parser
issues warnings when it detects any errors, some of which may alert the
user to a problem.  There should be few or no warnings when parsing, and
when finished parsing the state table should only contain entries for mails
which had yet to be delivered when the log files ended, or were accepted
before the log files began.  Verifying by inspection that the parser
correctly deals with all the mails in the test log files is infeasible, but
verifying a subset of those mails is a tractable, if extremely time
consuming, task.  A sample of 6000 log lines was parsed with all debugging
options enabled, resulting in 167,448 lines of output.  All 167,448 lines
were examined in conjunction with the input log file and a dump of the
resulting database, verifying that for each of the input lines the parser
used the correct rule and executed the correct action, which in turn
produced the correct result and inserted the correct data in the database.
The log segment produced 4 warnings (about mails which had started before
the log file segment), 10 mails remaining in the state tables, and 1625
connections correctly entered in the database.  Given the evidence detailed
above, the author is confident that the false positive rate when
reconstructing a mail is zero.

XXX DO I NEED THIS PARAGRAPH\@?
Parser coverage is divided into two topics in this section: log lines
covered and mails covered.  The former is initially more important, as the
parser must successfully parse every line if it is to be complete, but
subsequently the latter takes precedence because reproducing the path a
mail takes through Postfix is the aim of the parser.  Increasing the
percentage of log lines parsed is relatively simple and non-intrusive:
adding new rules or modifying existing rules is simplified by the
separation of rules, actions and framework.  Improving the logical coverage
is harder, as the actions taken by Postfix must be reconstructed by the
author, and the new sequence of actions integrated into the existing model
without breaking the existing parsing.  Detecting a deficiency in the
parsing algorithm is also significantly harder than detecting unparsed log
lines, as the parser will warn about any unparsed line, whereas discovering
a flaw in the parser requires understanding of the warnings produced and
the mails remaining in the state table.  Rectifying a flaw in the parser
requires an understanding of both the parser and Postfix's log files, and
investigative work to determine the cause of the deficiency, followed by
further examination of the log files in developing a solution.
